On the Dynamics of Anthropomorphism in Robotics

### 2013-07-11 dynamics-anthropomorphism.tex
refinements in the first sections

\abstract{While anthropomorphism in robotics is a commonly discussed trait of
HRI, it is paradoxically an overlooked research topic. This article
attempts first at providing a comprehensive synthesis of the social
phenomenon of \textit{perceiving human-like characteristics in non-human
agents and attributing those characteristics to robots}. We draw on
social sciences and psychology, as well as on a critical survey of
existing literature in the HRI community, to ground the concept of
\textit{anthropomorphism}. We further briefly present different kinds
of \textit{anthropomorphic design} of robots and outline their role in
social robotics\fxfatal{do we really do?}. We then suggest to go beyond the traditional
perception of anthropomorphism as a static feature of a system: we propose to
understand anthropomorphism as a dynamic, non-monotonic and
context-dependent process, which evolves over time and accounts for
special events like the so-called \textit{novelty effect}. To this end,
we introduce a model of anthropomorphism that analyzes the phenomenon
along three interaction phases. These findings are supported by a long-term
study conducted in a real human environment.}

@@

The essence of anthropomorphism is perceiving human-like characteristics in either real or imagined non-human agents. According to Epley \textit{et al.} \cite{epley_when_2008}, these human-like characteristics may include physical appearance (such as a religious agent believed to look human-like), emotional states perceived to be uniquely human, or inner mental states and motivations. Real or imagined non-human agents can be "anything that acts--or is believed to act--with apparent independence, including non-human animals, natural forces, religious agents, technological gadgets, or mechanical devices". Such anthropomorphic representations are important determinants of how a person behaves towards these agents, or how a person may behave \fxnote*{in light?}{in light} of these agents.

@@

In this context, anthropomorphism has received \fxnote*{contradict with intro}{remarkable attention} in the study of humans interacting with artificial entities, and has been a subject in Human-Computer Interaction (HCI) and Human-Robot Interaction (HRI). Following the \textit{Media Equation}, it has been described that people mindlessly react socially to computers \cite{reeves_media_1996}. The phenomenon has also been observed with virtual characters \hl{cite [50, 100] in rosenthal-von der P{\"u}tten (2013)}, and more recent studies investigated how far people ascribe personality or intentions to artificial agents \hl{references}\fxnote{Rosenthal is not enough?} and whether they show empathetic behavior toward robots \cite{rosenthal-vonderputten_experimental_2013}. The common base for this research is the assumption that humans perceive or react to a non-human agent or artifact as somewhat similar to a human.

We furthermore use in this article \emph{anthropomorphic effects} to denote the
observable correlates of anthropomorphism. These include, amongst others,
verbal behaviors (like the use of pronouns ``he/she''), kinesics (like hugging
the robot) or anticipation of certain robot's actions.

@@

However, not everything colloquially labeled as \emph{anthropomorphism} really is anthropomorphism. Different, even partially conflicting, understandings of anthropomorphism exist across and within disciplines \cite{duffy_anthropomorphism_2002}. Epley \textit{et al.} \cite{epley_when_2008} try to narrow down and clarify the understanding of anthropomorphism in psychology by explaining what it is \textit{not}. The authors list four things that anthropomorphism does not include: First, anthropomorphism does not include behavioral descriptions of observable actions. Anthropomorphism requires going beyond what is directly observable. Second, anthropomorphism does not merely entail animism, as animate life is not a uniquely human-like characteristic. Third, anthropomorphism does not include any requirement of reasoned or reflective endorsement of an inference (\hl{the strength of anthropomorphic inferences can vary} \footnote{for further explanation the reader may refer the original work \cite{epley_when_2008}.}).	Finally, anthropomorphism is not necessarily inaccurate. A psychological theory should take this (in)accuracy in anthropomorphism into account and predict variability in the tendency to perceive human-like traits in non-human agents\fxfatal{End of paragraph is unclear. We need to reformulate}.

%ANTHROPOMOPRHIC DESIGN

Anthropomorphism represents just one of many examples of induction whereby "people reason about an unknown stimulus based on a better-known representation of a related stimulus" \cite{epley_when_2008}, in this case reasoning about a non-human agent based on representation of the self or humans. This supports the usual role we assign to anthropomorphism in robotics: the anthropomorphic design of a robot encourages people to reason about this unfamiliar system \emph{as if} it was a human. To increase this tendency, robots are designed with anthropomorphic forms (human-like forms or lifelike forms, in the broader sense), or are able to display and use human-social behavior.

\paragraph{Anthropomorphism in robotics} Literature on anthropomorphism in robotics is quite diverse and one strives hard to extract a coherent conclusion. There is no commonly accepted definition about the terms used and the terminology is not clear. However the terms \textit{anthropomorphic or human-like are often used as if their meanings are clear and agreed upon"} \cite{persson_anthropomorphism_2000}. Consequently, Duffy \cite{duffy_anthropomorphism_2002} argues that terms might even be misused. For instance, some researchers refer to \textit{"the robot's level of anthropomorphism"} \cite{bartneck_is_2007}, however, others argue that anthropomorphism emerges in the \textit{interaction} between the technology and the user, thus a system or an artifact does not "contain anthropomorphism" \textit{per se} \cite{persson_anthropomorphism_2000} but only gives rise to the process of anthropomorphizing in a given user and situation. Further, robotics is a multi-disciplinary field and researchers from very different domains might have diverse or even contradictory understandings of anthropomorphism. Epley \textit{et al.} \cite{epley_seeing_2007} notice that despite the fact that anthropomorphism in the perception of non-human agents is commonly observed, it is poorly understood. The phenomenon itself has been found to be very complex but sometimes subtle and hard to study. Another constraint that limits the reliability of some findings is, that most experiments on anthropomorphism are conducted in controlled lab-settings during short-term interactions. This experimental setting (short term and laboratory context) can be critical when studying a social phenomenon like anthropomorphism and possibly lead to over-interpretations. Some studies also seem to contrast two or more different anthropomorphic and mechanical systems that might not be appropriate for comparison under the given context and research question. \hl{say more}

>>>>

\abstract{While anthropomorphism in robotics is a commonly discussed trait of
HRI, it paradoxically lacks formal grounds. This article
attempts first at providing a comprehensive synthesis of the social
phenomenon of \textit{perceiving human-like characteristics in non-human
agents and attributing those characteristics to robots}. We draw on
social sciences and psychology, as well as on a critical survey of
existing literature in the HRI community, to ground the concept of
\textit{anthropomorphism}. We then suggest to go beyond the traditional
perception of anthropomorphism as a static feature of a system: we propose to
understand anthropomorphism as a dynamic, non-monotonic and
context-dependent process, which evolves over time and accounts for
special events such as the so-called \textit{novelty effect}. To this end,
we introduce a model of anthropomorphism that analyzes the phenomenon
along three interaction phases. These findings are supported by a long-term
study conducted in a real human environment.}

@@

The essence of anthropomorphism is the perception of human-like characteristics in either real or imagined non-human agents. According to Epley \textit{et al.} \cite{epley_when_2008}, these human-like characteristics may include physical appearance, emotional states perceived to be uniquely human, or inner mental states and motivations. Real or imagined non-human agents can be "anything that acts -- or is believed to act -- with apparent independence, including non-human animals, natural forces, religious agents, technological gadgets, or mechanical devices". Such anthropomorphic representations are important determinants of how a person perceives and in turn behaves towards these agents, or how a person may behave in light of these agents \cite{epley_when_2008}.

@@

Phenomenons related to anthropomorphism -- along with questions such as "how much human do we perceive in the non-human" or "how human-like do we treat non-human artifacts" (and why?) -- have received remarkable attention in the domains of Human-Computer Interaction (HCI) and Human-Robot Interaction (HRI): the study of humans interacting with an artificial system / machine. Following the \textit{Media Equation}, it has been described that people mindlessly react socially to computers \cite{reeves_media_1996} and may even attribute human (social) characteristics to their artificial counterpart. The phenomenon has also been observed with virtual characters \hl{cite [50, 100] in rosenthal-von der P{\"u}tten (2013)}, and more recent studies investigated how far people ascribe personality or intentions to artificial emobdied agents or whether they show empathetic behavior toward robots \cite{rosenthal-vonderputten_experimental_2013}. The common base for this research is the assumption that humans perceive or react to a non-human agent or artifact as somewhat similar to a human.

In this article we will use the term \emph{anthropomorphic effect} to denote observable correlates of anthropomorphism. With these we include, amongst others,
verbal utterances (such as direct speech or the use of pronouns "he/she"), (social) gestures (such as a hug, waving, deictic gestures) or \hl{anticipation of certain actions of an agent [what??]}.

@@

However, not everything colloquially labeled as \emph{anthropomorphism} really is anthropomorphism. Different, even partially conflicting, understandings of anthropomorphism exist across and within disciplines \cite{duffy_anthropomorphism_2002}. Epley \textit{et al.} \cite{epley_when_2008} try to narrow down and clarify the understanding of anthropomorphism in psychology by explaining what it is \textit{not}. The authors list four things that anthropomorphism does not include: First, anthropomorphism does not include behavioral descriptions of observable actions. Anthropomorphism requires going beyond what is directly observable. Second, anthropomorphism does not merely entail animism, as animate life is not a uniquely human-like characteristic. Third, anthropomorphism does not include any requirement of reasoned or reflective endorsement of an inference (\hl{the strength of anthropomorphic inferences can vary} \footnote{for further explanation the reader may refer to the original work \cite{epley_when_2008}.}).	Finally, anthropomorphism is not necessarily inaccurate. A psychological theory should take this (in)accuracy in anthropomorphism into account and predict variability in the tendency to perceive human-like traits in non-human agents.
In conclusion, it has been noted that anthropomorphism is a very complex phenomenon - in itself, its understanding, and when studying it. A common formal ground is missing; however, it is important to be clear about what exactly we mean when using the term anthropomorphism.

\paragraph{Anthropomorphism in robotics \hl{[I think this first paragraph should go somewhere else, later probably]}}

Anthropomorphism represents just one of many examples of induction whereby "people reason about an unknown stimulus based on a better-known representation of a related stimulus" \cite{epley_when_2008}, in this case reasoning about a non-human agent based on representation of the self or other humans. This supports the usual role we assign to anthropomorphism in robotics: the anthropomorphic design of a robot encourages people to reason about this unfamiliar system \emph{as if} it was a human. To increase this tendency, robots are designed with anthropomorphic forms (human-like or lifelike forms, in the broader sense), or are able to display and use human-social cues when interacting with humans.\\

Literature on anthropomorphism in robotics is quite diverse and one strives hard to extract a coherent conclusion. One reason for this might be that robotics is a multi-disciplinary field and researchers from very different domains might integrate diverse or even contradictory understandings of the phenomenon. Despite the fact that there is no commonly accepted definition of anthropomorphism in robotics, and also the terminology is not clear, the terms \textit{anthropomorphic} or \textit{human-like} are often used as if their meanings were clear and agreed upon \cite{persson_anthropomorphism_2000}. Duffy \cite{duffy_anthropomorphism_2002} even argues that the terms might be misused. For instance, some researchers refer to \textit{"the robot's level of anthropomorphism"} \cite{bartneck_is_2007}, whereas others argue that the system or artifact itself does not "contain anthropomorphism" \textit{per se} but only gives rise to the process of anthropomorphizing in a given user and situation. Consequently, anthropomorphism emerges in the \textit{interaction} between the technology and the user \cite{persson_anthropomorphism_2000}.

Paradoxically, anthropomorphism in the perception of non-human agents (such as robots) is commonly observed but at the same time poorly understood \cite{epley_seeing_2007}. Why? The phenomenon itself has been found to be very complex. Still, it can sometimes also be subtle and as such hard to study. What might also account for the rather poor (or controversial) understanding of anthropomorphism in robotics, is the limited reliability of some findings (or interpretations) from experiments investigating anthropomorphism. Most of these experiments are conducted in controlled lab-settings during short-term human-robot interactions. This experimental setting (short term and laboratory context) can be critical when studying a social phenomenon like anthropomorphism and possibly lead to over-interpretations. Some studies also seem to contrast two or more different anthropomorphic and mechanical systems that might not be appropriate for comparison under the given context and research question. \hl{say more}
